---
title: "ChIPSeq MiSeq QC"
output:
  pdf_document: default
date: '`r format(Sys.Date(), "%B %d, %Y")`'
params:
  input: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(hms)
library(tidyverse)
library(scales)
library(knitr)

df <- read.table(
file = params$input,
sep="\t",
header=T,
stringsAsFactors = TRUE
)


parsed_input <- strsplit(basename(params$input), "_")

ssf <- parsed_input[[1]][1]

samples <- df %>% filter(Ctrl=="no")
```


## Sample representation in ``r basename(params$input)``

Batch size: `r nrow(df)`  
&#35; samples in SSF: `r df %>% select(Sample, Ctrl) %>% filter(Ctrl=="no") %>% nrow()`  
&#35; controls in SSF: `r df %>% select(Sample, Ctrl) %>% filter(Ctrl=="yes") %>% nrow()`  

## H3K27ac FRiP stratification (samples only) 

```{r stratified_FRiP, echo=FALSE}
df2 <- df %>% mutate(fcat=cut(mapq1PE_FRIP, breaks=c(-Inf, 0.03, 0.05, 0.1, Inf), labels=c("<3%","3%<X<5%", "5%<X<10%",">10%")))

kable(df2 %>% filter(Ctrl=="no") %>% group_by(fcat) %>% summarize (n = n()) %>% mutate(pct_fcat = percent(n/sum(n))))
```

## Results from Controls

```{r controls, echo=FALSE}

kable(df %>% filter(Ctrl=="yes") %>% mutate(H3K27ac_FRiP = percent(mapq1PE_FRIP)) %>% select(Sample, H3K27ac_FRiP, Tot_Reads))
```

## Evaluation of total reads 

Total reads: `r format(sum(df$Tot_Reads), big.mark = ",")`  
Avg reads (all): `r format(round(mean(df$Tot_Reads)), scientific=FALSE, big.mark = ",")`  
Avg reads (samples only): `r format(round(df %>% filter(Ctrl=="no")  %>% select(Sample, Tot_Reads) %>%  summarize(mean(Tot_Reads))), scientific=FALSE, big.mark = ",")`  


## Distribution of Total Reads

```{r Tot_Reads_plot, echo=FALSE}
plot(ggplot(data=df, aes(x = reorder(Sample, Tot_Reads), y=Tot_Reads)) + geom_bar(stat="identity", fill = "grey", colour = "black") + 
    geom_bar(data = df, aes(x = reorder(Sample, MAPQ1_Reads), y = MAPQ1_Reads), stat = "identity", fill = "grey", colour = "blue") +
    theme(axis.text.x = element_text(size=5, angle = 90, hjust = 1)) )
```



## Duplication rate

Average duplication rate (all): `r percent(mean(df$pDUPLICATION))`  
sample only duplication rate: `r percent(mean(samples$pDUPLICATION))`  


&#35; samples with >0.5% duplication: `r df %>% select(Sample, pDUPLICATION) %>% filter(pDUPLICATION>=0.005) %>% nrow()`  
&#35; samples with >1% duplication: `r df %>% select(Sample, pDUPLICATION) %>% filter(pDUPLICATION>=0.01) %>% nrow()`  

samples with >0.5% duplication:
```{r DUP_gt_1pct, echo=FALSE}
df %>% select(Sample, pDUPLICATION) %>% filter(pDUPLICATION > 0.005) %>% arrange(desc(pDUPLICATION)) %>% mutate(pDUP = percent(pDUPLICATION)) %>% select(Sample, pDUP) 
``` 

## Adapter rate (R1)

Average adapter rate (all): `r percent(mean(df$R1_ADAPTER))`  
sample-only adapter rate: `r percent(mean(samples$R1_ADAPTER))`  
&#35; samples with > 2% adapter: `r df %>% select(Sample, R1_ADAPTER) %>% filter(R1_ADAPTER>=0.02) %>% nrow()`  
&#35; samples with > 5% adapter: `r df %>% select(Sample, R1_ADAPTER) %>% filter(R1_ADAPTER>=0.05) %>% nrow()`  

samples with >2% adapter:
```{r , echo=FALSE}
df %>% select(Sample, R1_ADAPTER) %>% filter(R1_ADAPTER > 0.02) %>% arrange(desc(R1_ADAPTER)) %>% mutate(Adapter = percent(R1_ADAPTER)) %>% select(Sample, Adapter) 
```


