---
title: "ChIPSeq HiSeq QC"
output:
    pdf_document: default
date: '`r format(Sys.Date(), "%B %d, %Y")`'
params:
  input: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(hms)
library(tidyverse)
library(scales)
library(knitr)

df <- read.table(
file = params$input,
sep="\t",
header=T,
stringsAsFactors = TRUE
)


parsed_input <- strsplit(basename(params$input), "_")

ssf <- parsed_input[[1]][1]

samples <- df %>% filter(Ctrl=="no")
```


## Sample representation in ``r basename(params$input)``

Batch size: `r nrow(df)`  
&#35; samples in SSF: `r df %>% select(Sample, Ctrl) %>% filter(Ctrl=="no") %>% nrow()`  
&#35; controls in SSF: `r df %>% select(Sample, Ctrl) %>% filter(Ctrl=="yes") %>% nrow()`  

## H3K27ac FRiP stratification (samples only) 

```{r stratified_FRiP, echo=FALSE}
df2 <- df %>% mutate(fcat=cut(mapq1PE_FRIP, breaks=c(-Inf, 0.03, 0.05, 0.1, Inf), labels=c("<3%","3%<X<5%", "5%<X<10%",">10%")))

kable(df2 %>% filter(Ctrl=="no") %>% group_by(fcat) %>% summarize (n = n()) %>% mutate(pct_fcat = percent(n/sum(n))))
```

## Mean values of key metrics (samples only)

```{r MetricsMeans, echo=FALSE}
means_df <- df %>% filter(Ctrl=="no") %>% select(pctAnalyzed,mapq1PE_FRIP,R1_ADAPTER,pDUPLICATION,RSC) %>% summarize_if(is.numeric,funs(mean)) %>% transmute(H3K27ac_FRIP = percent(mapq1PE_FRIP) ,R1_ADAPTER = percent(R1_ADAPTER) ,pDUPLICATION = percent(pDUPLICATION), pctAnalyzed = percent(pctAnalyzed) ,RSC = RSC )

kable(means_df, row.names = FALSE)
```


```{r check_controls, echo=FALSE}

ctrl_num <- df %>% select(Sample, Ctrl) %>% filter(Ctrl=="yes") %>% nrow()
if ( ctrl_num==0) {
  show_text <- FALSE
} else {
  show_text <- TRUE
}

```

```{r controls, echo=FALSE, results='asis', eval=show_text}
cat("## Results from Controls")

kable(df %>% filter(Ctrl=="yes") %>% mutate(H3K27ac_FRiP = percent(mapq1PE_FRIP)) %>% select(Sample, H3K27ac_FRiP, Tot_Reads))
```

## Evaluation of total reads 

Total reads: `r format(sum(df$Tot_Reads), big.mark = ",")`  
Avg reads (all): `r format(round(mean(df$Tot_Reads)), scientific=FALSE, big.mark = ",")`  
Avg reads (samples only): `r format(round(df %>% filter(Ctrl=="no")  %>% select(Sample, Tot_Reads) %>%  summarize(mean(Tot_Reads))), scientific=FALSE, big.mark = ",")`  
Avg mapq1 reads (samples only): `r format(round(df %>% filter(Ctrl=="no")  %>% select(Sample, MAPQ1_Reads) %>%  summarize(mean(MAPQ1_Reads))), scientific=FALSE, big.mark = ",")` 

## Distribution of Total Reads

```{r Tot_Reads_plot, echo=FALSE}
ggplot(data=df, aes(x = reorder(Sample, Tot_Reads), y=Tot_Reads)) + geom_bar(stat="identity", fill = "grey", colour = "black") + 
    geom_bar(data = df, aes(x = reorder(Sample, MAPQ1_Reads), y = MAPQ1_Reads), stat = "identity", fill = "grey", colour = "blue") +
    theme(axis.text.x = element_text(size=5, angle = 90, hjust = 1)) + 
    geom_hline(yintercept=10000000, colour ="red")
```

```{r check_coverage_assess, echo=FALSE}

ctrl_num <- df %>% select(Sample, Ctrl) %>% filter(Ctrl=="yes") %>% nrow()
if ( mean(df$Tot_Reads)<10000000) {
  coverage_assess <- FALSE
} else {
  coverage_assess <- TRUE
}

```

```{r coverage_assess, echo=FALSE, eval=coverage_assess}
cat("samples with less than 10M reads after mapq1 filtering")
df %>% select(Sample, MAPQ1_Reads) %>% filter(MAPQ1_Reads < 10000000) %>% arrange(desc(MAPQ1_Reads)) %>% select(Sample, MAPQ1_Reads)

```


```{r check_excess_reads, echo=FALSE}

outlier_samples <- df %>% filter(Tot_Reads > mean(df$Tot_Reads) + 2* sd(df$Tot_Reads)) %>% nrow()
if ( outlier_samples==0) {
  excess_reads <- FALSE
} else {
  excess_reads <- TRUE
}

```

```{r excess_reads, echo=FALSE, eval=excess_reads}
cat("jackpot samples")
df %>% filter(Tot_Reads > mean(df$Tot_Reads) + 2* sd(df$Tot_Reads)) %>% select(Sample,Tot_Reads)

lost_reads <- df %>% filter(Tot_Reads > mean(df$Tot_Reads) + 2* sd(df$Tot_Reads)) %>% mutate(lost = Tot_Reads-(mean(df$Tot_Reads))) %>% select(Sample,lost) %>% summarize(sum(lost))

percent_excess <- toString(round(lost_reads/sum(df$Tot_Reads)*100, digits=1))
cat(percent_excess, '% of Hiseq run are excess reads for jackpot samples', sep = "")

```



## Duplication rate

Average duplication rate (all): `r percent(mean(df$pDUPLICATION))`  
sample only duplication rate: `r percent(mean(samples$pDUPLICATION))`  


&#35; samples with >3% duplication: `r df %>% select(Sample, pDUPLICATION) %>% filter(pDUPLICATION>=0.03) %>% nrow()`  
&#35; samples with >5% duplication: `r df %>% select(Sample, pDUPLICATION) %>% filter(pDUPLICATION>=0.05) %>% nrow()`  

samples with >0.5% duplication:
```{r DUP_gt_1pct, echo=FALSE}
df %>% select(Sample, pDUPLICATION) %>% filter(pDUPLICATION > 0.05) %>% arrange(desc(pDUPLICATION)) %>% mutate(pDUP = percent(pDUPLICATION)) %>% select(Sample, pDUP) 
``` 

## Adapter rate (R1)

Average adapter rate (all): `r percent(mean(df$R1_ADAPTER))`  
sample-only adapter rate: `r percent(mean(samples$R1_ADAPTER))`  
&#35; samples with > 2% adapter: `r df %>% select(Sample, R1_ADAPTER) %>% filter(R1_ADAPTER>=0.02) %>% nrow()`  
&#35; samples with > 5% adapter: `r df %>% select(Sample, R1_ADAPTER) %>% filter(R1_ADAPTER>=0.05) %>% nrow()`  

samples with >2% adapter:
```{r , echo=FALSE}
df %>% select(Sample, R1_ADAPTER) %>% filter(R1_ADAPTER > 0.02) %>% arrange(desc(R1_ADAPTER)) %>% mutate(Adapter = percent(R1_ADAPTER)) %>% select(Sample, Adapter) 
```


